---
title: "Vast Challenge 2024: Question 4"
author: "Ke Ke"
date: "18 May 2024"
date-modified: "last-modified"
format:
  html:
    code-fold: true
    code-tools: true
execute:
  warning: false
  freeze: true
---

## Getting Started

```{r}
pacman::p_load(tidytext, readtext, quanteda, tidyverse, jsonlite, igraph, tidygraph, ggraph, visNetwork, clock, graphlayouts,ggplot2)
```

## **Importing JSON File**

Direct import of the mc3.json file shows an error message indicating that there's an invalid character in the JSON text, specifically "NaN". As "NaN" is not recognised as a valid value, preprocessing of the JSON file to replace "NaN" is required.

```{r}
# Read the JSON file as text
json_text <- readLines("data/mc3.json" ,warn = FALSE)

# Replace "NaN" with "null"
json_text_fixed <- gsub("NaN", "null", json_text)

# Write the fixed JSON text back to a file
writeLines(json_text_fixed, "data/mc3_fixed.json")
```

Importing preprocessed mc3_fixed.json file

```{r}
mc3_data <- fromJSON("data/mc3_fixed.json")

```

## Data Cleaning

### Missing Values

Identify the percentage of missing values within the dataset

```{r}
# Function to calculate missing value percentages
calculate_missing_percentage <- function(df) {
  total_values <- nrow(df) * ncol(df)
  missing_values <- sum(is.na(df))
  missing_percentage <- (missing_values / total_values) * 100
  return(missing_percentage)
}
```

Missing percentage of nodes

```{r}
nodes_missing_percentage <- calculate_missing_percentage(mc3_data[["nodes"]])
nodes_missing_percentage
```

```{r}
nodes_missing_by_column <- sapply(mc3_data[["nodes"]], function(x) sum(is.na(x)) / length(x) * 100)
```

Missing percentage of edges

```{r}
links_missing_percentage <- calculate_missing_percentage(mc3_data[["links"]])
links_missing_percentage

links_missing_by_column <- sapply(mc3_data[["links"]], function(x) sum(is.na(x)) / length(x) * 100)
links_missing_by_column
```

Print missing percentages

```{r}
# 
print(nodes_missing_percentage)
print(nodes_missing_by_column)
print(links_missing_percentage)
print(links_missing_by_column)
```

::: panel-tabset
## Observations:

-   **Nodes Data:**

    -   `ProductServices`, `PointOfContact`, `HeadOfOrg`, `founding_date`, `revenue`, and `TradeDescription` columns have a high percentage of missing values (around 85%).

    -   The `dob` column has about 14.7% missing values.

    -   Other columns (`type`, `country`, `_last_edited_by`, `_last_edited_date`, `_date_added`, `_raw_source`, `_algorithm`, and `id`) have no missing values.

-   **Links Data:**

    -   `end_date` has a very high percentage of missing values (around 99.5%).

## Actions:

-   Filled missing values in `HeadOfOrg` with "Unknown".

-   Filled missing values in `revenue` with 0.

-   Filled missing values in `start_date` and `end_date` with "Unknown".
:::

Handle missing values

```{r}


# Select crucial columns and fill missing values where appropriate
cleaned_nodes <- mc3_data[["nodes"]] %>%
  select(id, type, country, HeadOfOrg, revenue,ProductServices,PointOfContact,founding_date,TradeDescription,dob,
         `_last_edited_by`, `_last_edited_date`, `_date_added`, `_raw_source`, `_algorithm`) %>%
  mutate(HeadOfOrg = ifelse(is.na(HeadOfOrg), "Unknown", HeadOfOrg),
         revenue = ifelse(is.na(revenue), 0, revenue))

# Handle missing values in links
# Select crucial columns and fill missing values where appropriate
cleaned_links <- mc3_data[["links"]] %>%
  select(key,source, target, type, start_date, end_date, `_last_edited_by`, `_last_edited_date`, `_date_added`, `_raw_source`, `_algorithm`) %>%
  mutate(start_date = ifelse(is.na(start_date), "Unknown", start_date),
         end_date = ifelse(is.na(end_date), "Unknown", end_date))

# Ensure proper data types
cleaned_nodes <- cleaned_nodes %>%
  mutate(
    id = as.character(id),
    type = as.character(type),
    country = as.character(country),
    HeadOfOrg = as.character(HeadOfOrg),
    revenue = as.numeric(revenue),
    `_last_edited_by` = as.character(`_last_edited_by`),
    `_last_edited_date` = as.character(`_last_edited_date`),
    `_date_added` = as.character(`_date_added`),
    `_raw_source` = as.character(`_raw_source`),
    `_algorithm` = as.character(`_algorithm`)
  )

cleaned_links <- cleaned_links %>%
 mutate(
    source = as.character(source),
    target = as.character(target),
    type = as.character(type),
    start_date = as.character(start_date),
    end_date = as.character(end_date),
    `_last_edited_by` = as.character(`_last_edited_by`),
    `_last_edited_date` = as.character(`_last_edited_date`),
    `_date_added` = as.character(`_date_added`),
    `_raw_source` = as.character(`_raw_source`),
    `_algorithm` = as.character(`_algorithm`)
  )



```

### Check for data types

```{r}


# Ensure correct data types for nodes
cleaned_nodes <- cleaned_nodes %>%
  mutate(
    id = as.character(id),
    type = as.character(type),
    country = as.character(country),
    HeadOfOrg = as.character(HeadOfOrg),
    revenue = as.numeric(revenue),
      dob = as.POSIXct(dob, format="%Y-%m-%dT%H:%M:%S"),
    `_last_edited_by` = as.character(`_last_edited_by`),
    `_last_edited_date` = as.POSIXct(`_last_edited_date`, format="%Y-%m-%dT%H:%M:%S"),
    founding_date=as.POSIXct(founding_date, format="%Y-%m-%dT%H:%M:%S"),
    `_date_added` = as.POSIXct(`_date_added`, format="%Y-%m-%dT%H:%M:%S"),
    `_raw_source` = as.character(`_raw_source`),
    `_algorithm` = as.character(`_algorithm`)
    
  )

# Ensure correct data types for links
cleaned_links <- cleaned_links %>%
 mutate(
    source = as.character(source),
    target = as.character(target),
    type = as.character(type),
    start_date = as.POSIXct(start_date, format="%Y-%m-%dT%H:%M:%S"),
    end_date = as.POSIXct(end_date, format="%Y-%m-%dT%H:%M:%S"),
    `_last_edited_by` = as.character(`_last_edited_by`),
    `_last_edited_date` = as.POSIXct(`_last_edited_date`, format="%Y-%m-%dT%H:%M:%S"),
    `_date_added` = as.POSIXct(`_date_added`, format="%Y-%m-%dT%H:%M:%S"),
    `_raw_source` = as.character(`_raw_source`),
    `_algorithm` = as.character(`_algorithm`)
  )

# Print cleaned data for inspection
glimpse(cleaned_nodes)
glimpse(cleaned_links)


```

### **Changing field name**

```{r}
cleaned_nodes <- cleaned_nodes %>%
  rename("last_edited_by" = "_last_edited_by",
         "date_added" = "_date_added",
         "last_edited_date" = "_last_edited_date",
         "raw_source" = "_raw_source",
         "algorithm" = "_algorithm") 

cleaned_links<- cleaned_links %>%
  rename("last_edited_by" = "_last_edited_by",
         "date_added" = "_date_added",
         "last_edited_date" = "_last_edited_date",
         "raw_source" = "_raw_source",
         "algorithm" = "_algorithm") 
```

### Split *'type'* column into separate columns

We are going to tidy the type column by creating two columns "entity2,entity3".

```{r}
word_list1 <- strsplit(cleaned_nodes$type, "\\.")
max_elements1 <- max(lengths(word_list1))
word_list_padded1 <- lapply(word_list1, 
function(x) c(x, rep(NA, max_elements1 - length(x))))
word_df1 <- do.call(rbind, word_list_padded1)
colnames(word_df1) <- paste0("entity", 1:max_elements1)
word_df1 <- as_tibble(word_df1) %>%
  select(entity2, entity3)
class(word_df1)
```

The steps below will be used to split text in type column into two columns

```{r}
word_list <- strsplit(cleaned_links$type, "\\.")
max_elements <- max(lengths(word_list))
word_list_padded <- lapply(word_list, 
function(x) c(x, rep(NA, max_elements - length(x))))
word_df <- do.call(rbind, word_list_padded)
colnames(word_df) <- paste0("event", 1:max_elements)
word_df <- as_tibble(word_df) %>%
  select(event2, event3)
class(word_df)
```

Since the output above is a matrix, the code chunk above is used to convert word_df into a tibble data.frame.

```{r}
cleaned_nodes <- cleaned_nodes %>%
  cbind(word_df1)
```

```{r}
cleaned_links <- cleaned_links %>%
  cbind(word_df)
```

The code chunk above appends the extracted columns back to edges tibble data.frame.

```{r}
write_rds(cleaned_nodes, "data/rds/cleaned_nodes.rds")
write_rds(cleaned_links, "data/rds/cleaned_links.rds")
```

above code write into R **rds** file format.

## Question 3

By analyzing the ownership structure, we tracked changes in most influential individuals (VIP) networks over time, identifying key individuals with increasing influence.

### Part 1: Data Wrangling

Split the nodes into people and companies, and filter ownership-related edges

```{r}
# Split the nodes into people and companies
nodes_people <- cleaned_nodes %>% filter(entity2 == "Person")
nodes_company <- cleaned_nodes %>% filter(entity2 == "Organization")



```

```{r}

# Filter the links to include only ownership-related edges
links_owns <- cleaned_links %>% filter(event2 == "Owns")

```

```{r}
nodes_people <- nodes_people %>%
  rowwise() %>%
  mutate('no_owns' = sum(links_owns$source == id))

nodes_people$no_owns <- as.numeric(nodes_people$no_owns)

```

```{r}
# Calculate the unique counts of 'no_owns' and their corresponding counts and percentages
owns_summary <- nodes_people %>%
  group_by(no_owns) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)

# Display the summary
print(owns_summary)

```

To define and identify influential people based on an ownership threshold. It filters the nodes to keep only those with a significant number of ownerships

```{r}
# Define the threshold for 'influential'
vip_threshold <- 91

# Filter to keep only influential people and select relevant columns
vip <- nodes_people %>%
  filter(no_owns >= vip_threshold) %>%
  select(id, country, dob, last_edited_date, date_added, no_owns)

# Display the updated vip data frame
glimpse(vip)


```

filter the ownership connections to include only those involving these influential individuals

```{r}
# Filter links_owns to keep only those connections where the source is in the vip list
vip_connections <- links_owns %>%
  filter(source %in% vip$id)%>%
  select(source, target,start_date,end_date,last_edited_date, date_added)

# Display the updated vip_connections data frame
glimpse(vip_connections)

```

### Part 1: Network Graph

Finally, plot the network, highlighting the structure and connections of influential individuals.

```{r}
# Create graph from VIP connections
g_vip <- graph_from_data_frame(d = vip_connections, directed = TRUE)

# Identify VIPs and Companies
V(g_vip)$type <- ifelse(V(g_vip)$name %in% nodes_people$id, "VIP", "Company")

# Define colors and sizes
V(g_vip)$color <- ifelse(V(g_vip)$type == "VIP", "blue", "orange")
V(g_vip)$size <- ifelse(V(g_vip)$type == "VIP", 8, 5)

# Plot the network
plot(g_vip, vertex.label = NA, vertex.size = V(g_vip)$size, edge.arrow.size = 0.5, 
     vertex.color = V(g_vip)$color, main = "VIP Connections Network")
```

The plot represents the VIP Connections Network, with blue nodes indicating influential VIPs and orange nodes representing companies they own. Directed edges illustrate ownership, pointing from VIPs to companies. This visualization highlights the dense centrality of VIPs, showcasing their extensive control across multiple companies. By examining these connections, we can infer the structure and extent of VIP influence within the network and help FishEye identify influential individuals within the business network, highlighting ownership structures and central figures. By tracking ownership changes over time, FishEye can pinpoint who controls companies involved in illegal fishing activities.

While this plot provides a static snapshot, in the following we shall create similar plots for different time periods can reveal changes in ownership and influence over time.

### Part 2: Temporal Analysis

Aggregate Ownership Changes by Year

```{r}
change_over_time1 <- links_owns %>%
  group_by(start_date) %>%
  summarize(count = n()) %>%
  drop_na()

links_owns<- links_owns %>%
  mutate(start_year = format(start_date, "%Y"))

# Aggregate ownership changes by year
change_over_time <- links_owns %>%
  group_by(start_year) %>%
  summarize(count = n()) %>%
  drop_na()

```

Create plots to visualize the changes in ownership over time.

```{r}
# Plot changes over time
ggplot(change_over_time, aes(x = as.numeric(start_year), y = count)) +
  geom_line() +
  geom_point() +
  labs(title = "Changes in Ownership Over Time",
       x = "Year",
       y = "Number of Ownership Changes") +
  theme_minimal()

```

### Part 2: Network Graph by Year

Given the significant increase in data from 2004 onwards, focusing on every 10 years from 2005 to 2035 would provide a more detailed analysis of changes in ownership and influence.

```{r}
# Specify the year 
filter_year <- 2005

# Filter vip_connections by start_year
vip_connections_filtered <- vip_connections %>%
  filter(format(start_date, "%Y") == filter_year)

# Create the graph object from the filtered vip_connections
g_vip_filtered <- graph_from_data_frame(d = vip_connections_filtered, directed = TRUE)

# Identify VIPs (nodes_people) and Companies
V(g_vip_filtered)$type <- ifelse(V(g_vip_filtered)$name %in% nodes_people$id, "VIP", "Company")

# Define colors and sizes
V(g_vip_filtered)$color <- ifelse(V(g_vip_filtered)$type == "VIP", "blue", "orange")
V(g_vip_filtered)$size <- ifelse(V(g_vip_filtered)$type == "VIP", 8, 5)

# Plot the network
p2005<-plot(g_vip_filtered, vertex.label = NA, vertex.size = V(g_vip_filtered)$size, edge.arrow.size = 0.5, 
     vertex.color = V(g_vip_filtered)$color, main = paste("VIP Connections Network for", filter_year))

```

In 2005, the network shows a relatively sparse structure with a moderate number of connections. VIPs (blue nodes) are moderately interconnected, indicating a balanced distribution of influence among several key players.

```{r}
# Specify the year
filter_year <- 2015

# Filter vip_connections by start_year
vip_connections_filtered_2015 <- vip_connections %>%
  filter(format(start_date, "%Y") == filter_year)

# Create the graph object from the filtered vip_connections
g_vip_filtered_2015 <- graph_from_data_frame(d = vip_connections_filtered_2015, directed = TRUE)

# Identify VIPs (nodes_people) and Companies
V(g_vip_filtered_2015)$type <- ifelse(V(g_vip_filtered_2015)$name %in% nodes_people$id, "VIP", "Company")

# Define colors and sizes
V(g_vip_filtered_2015)$color <- ifelse(V(g_vip_filtered_2015)$type == "VIP", "blue", "orange")
V(g_vip_filtered_2015)$size <- ifelse(V(g_vip_filtered_2015)$type == "VIP", 8, 5)

# Plot the network
p2015 <- plot(g_vip_filtered_2015, vertex.label = NA, vertex.size = V(g_vip_filtered_2015)$size, edge.arrow.size = 0.5, 
     vertex.color = V(g_vip_filtered_2015)$color, main = paste("VIP Connections Network for", filter_year))



```

By 2015, the network has grown denser, suggesting increased interconnectedness and influence consolidation. More VIPs are connected to multiple companies (orange nodes), indicating a significant rise in their influence and control over the network.

```{r}

# Specify the year
filter_year <- 2025

# Filter vip_connections by start_year
vip_connections_filtered_2025 <- vip_connections %>%
  filter(format(start_date, "%Y") == filter_year)

# Create the graph object from the filtered vip_connections
g_vip_filtered_2025 <- graph_from_data_frame(d = vip_connections_filtered_2025, directed = TRUE)

# Identify VIPs (nodes_people) and Companies
V(g_vip_filtered_2025)$type <- ifelse(V(g_vip_filtered_2025)$name %in% nodes_people$id, "VIP", "Company")

# Define colors and sizes
V(g_vip_filtered_2025)$color <- ifelse(V(g_vip_filtered_2025)$type == "VIP", "blue", "orange")
V(g_vip_filtered_2025)$size <- ifelse(V(g_vip_filtered_2025)$type == "VIP", 8, 5)

# Plot the network
p2025 <- plot(g_vip_filtered_2025, vertex.label = NA, vertex.size = V(g_vip_filtered_2025)$size, edge.arrow.size = 0.5, 
     vertex.color = V(g_vip_filtered_2025)$color, main = paste("VIP Connections Network for", filter_year))

```

The network continues to expand in 2025, displaying even more complexity and interconnections. This period likely represents a peak in influence for several VIPs, with many of them owning shares in numerous companies, suggesting increased market control.

```{r}
# Specify the year
filter_year <- 2035

# Filter vip_connections by start_year
vip_connections_filtered_2035 <- vip_connections %>%
  filter(format(start_date, "%Y") == filter_year)

# Create the graph object from the filtered vip_connections
g_vip_filtered_2035 <- graph_from_data_frame(d = vip_connections_filtered_2035, directed = TRUE)

# Identify VIPs (nodes_people) and Companies
V(g_vip_filtered_2035)$type <- ifelse(V(g_vip_filtered_2035)$name %in% nodes_people$id, "VIP", "Company")

# Define colors and sizes
V(g_vip_filtered_2035)$color <- ifelse(V(g_vip_filtered_2035)$type == "VIP", "blue", "orange")
V(g_vip_filtered_2035)$size <- ifelse(V(g_vip_filtered_2035)$type == "VIP", 8, 5)

# Plot the network
p2035 <- plot(g_vip_filtered_2035, vertex.label = NA, vertex.size = V(g_vip_filtered_2035)$size, edge.arrow.size = 0.5, 
     vertex.color = V(g_vip_filtered_2035)$color, main = paste("VIP Connections Network for", filter_year))

```

In 2035, the network structure shifts to a star-like formation, where a central VIP appears to have gained substantial influence, with direct connections to numerous companies. This indicates a significant consolidation of power and influence, where a few key players dominate the network.

::: callout-caution
Initially, influence is distributed among several key players, but over the years, it becomes concentrated among fewer individuals, leading to a highly centralized network by 2035. This centralization of power can be both an opportunity for streamlined decision-making and a risk for monopolistic control. Monitoring these changes is crucial for regulatory bodies like FishEye to ensure fair practices and prevent illegal activities within the network.
:::
